---
title: ''
author: "Cori Mar"
date: "11/30/2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Using the Human Mortality Data

1. Go to www.mortality.org
2. Create a user name and password
3. Login
4. Choose country: Australia
5. Choose data: Births
6. Download file: Australia_Births_Raw.txt
7. Strip header from file and save: Australia_Births.dat

```{r}
### read data file
dd <- read.table(file="Australia_Births.dat", header=TRUE)

```

#### Plot Australia births

```{r, echo=FALSE}

### plot total
plot(dd$Year, dd$Total/1000, ylim=c(0,325), type="l", las=1,
     xlab="Year", ylab="Thousands",main="Australia: Births")

### grid lines
segments(x0=seq(1860, 2000, 20), x1=seq(1860, 2000, 20), y0=-10, y1=500, lty="dotted")
segments(y0=seq(0, 3000, 50), y1=seq(0, 3000, 50), x0=-10, x1=3000, lty="dotted")

### plot females
lines(dd$Year, dd$Female/1000, col="red")
### plot males
lines(dd$Year, dd$Male/1000, col="blue")

legend(1860,300, legend=c("Total","Male","Female"), lty=1, col=c("black","blue","red"))

```

#### Read table of Australia Life expectancy at birth

```{r}
dd.le <- read.table(file="Australia_LifeExpBirth_1x10.dat", header = TRUE)
```
  
  
#### Print table of Australia Life expectancy at birth

##### Raw output

```{r}
dd.le
```

##### Raw output without hashtags using comment=''

```{r comment=''}
dd.le
```

##### Prettier output using kable
```{r results='asis'}
library(knitr)
kable(dd.le, caption="Australia Life Expectancy at Birth")
```

#### Regression 
```{r}
# regression births as a function of year
fit <- lm(Total ~ Year, data=dd)
```

#### Regression output using kable

```{r results='asis'}
test <- summary(fit)
kable(test$coefficients, caption="Regression Table")
```

#### Text with imbedded R

The number of live births in Australia in `r dd$Year[1]` was `r dd$Total[1]`.  By `r dd$Year[nrow(dd)]`, this number was `r dd$Total[nrow(dd)]`.

#### Text with imbedded R with numbers formatted with commas

The number of live births in Australia in `r dd$Year[1]` was `r prettyNum(dd$Total[1],big.mark=",")`.  By `r dd$Year[nrow(dd)]`, this number was `r prettyNum(dd$Total[nrow(dd)],big.mark=",")`.


